ICM testing

### 2023-04-05 Scripts/ASR_stocks.Rmd

```{r, prep-data,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# OK, so using the ICES assessments here's what we get for North Sea cod.
library(readxl)
library(tidyverse)
library(rio)
library(ggthemes)
library(cowplot)

# # OK, lets give this a spin...
# funs <- c("https://raw.githubusercontent.com/Dave-Keith/ICM/main/Scripts/functions/backwards_sim.r")
# # Now run through a quick loop to load each one, just be sure that your working directory is read/write!
# for(fun in funs) 
# {
#   download.file(fun,destfile = basename(fun))
#   source(paste0(getwd(),"/",basename(fun)))
#   file.remove(paste0(getwd(),"/",basename(fun)))
# }

source("D:/Github/ICM/Scripts/functions/backwards_sim.R")

# Choose 5 ICES stocks that we have the necessary data for
#ASR <- read_xlsx("../Data/ASR_2018.xlsx", sheet = "ICES")
ASR1 <- read.csv("D:/Github/ICM/Data/ICM_data_NE_atlantic_stocks.csv")
ASR <- read.csv("D:/Github/ICM/Data/ICM_data_NW_atlantic_stocks.csv")
ASR <- rbind(ASR1,ASR)
datatypes <- unique(gsub(x = names(ASR), pattern = "[^a-zA-Z]", replacement=""))

# Replace 0's in Num.tot with NA so the rest of this works as Freya designed it to...
ASR$Num.tot[ASR$Num.tot == 0] <- NA
# we want:
# Year, Num, WA, Catch, AM, NM, StockID, Management, Area, Order, Family, Genus, Species
ASRdat <- ASR[,c(grep(x=names(ASR), "Num"),
                 grep(x=names(ASR), "WA"),
                 grep(x=names(ASR), "Catch"),
                 grep(x=names(ASR), "AM"),
                 grep(x=names(ASR), "NM"))]
# Making all the data numeric that should be numeric
ASRdat <- apply(X = ASRdat, 2, as.numeric)
# Getting the species info back
ASRsp <- ASR[, which(!1:length(names(ASR)) %in% grep(x=names(ASR), ".", fixed=T))]
# And binding it all back together
ASR_trim <- cbind(ASRsp, ASRdat)

# need a unique ID for stock
#table(ASR_trim$Management, ASR_trim$Species)
ASR_trim$Stock <- paste0(ASR_trim$Management, "_", ASR_trim$Area, "_", ASR_trim$Genus, "_", ASR_trim$Species)

@@

@@

# Stocks <- ASR_stocks %>%
#                       dplyr::group_by(Stock, Species, type) %>%
#                       dplyr::summarize(ages=length(unique(age)),
#                                 years=length(unique(Year))) %>%
#                       dplyr::arrange(-years, -ages) %>%
#                       dplyr::filter(!Species=="morhua") %>%
#                       dplyr::distinct(Stock) %>%
#                       dplyr::pull(Stock)
# 
# print(Stocks)
Stocks <- Stocks$Stock
# Going to remove Sebastes norvegicus because we can't get fecundity as the only data we have is total numbers.
#Stocks <- Stocks[Stocks != "ICES-AFWG_DEEP1-2_Sebastes_norvegicus"]
# Esmarki doesn't work great as it is assessed using a quarterly modle.
#Stocks <- Stocks[Stocks != "ICES-WGNSSK_NS 4-3aN_Trisopterus_esmarkii"]

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  print(i)
  ASR_sub <- ASR_long %>%
    dplyr::select(Year, Stock, type, age, value) %>%
    dplyr::filter(Stock==i)

@@

  # For the years of mx without SSB info to inform them
  if(minage >0 ) 
  {
    mx.fill <- as.data.frame(matrix(rep(colMeans(mx,na.rm=T),minage),nrow=minage,byrow=T),colnames = names(mx))
    names(mx) <- names(mx.fill)
   # mx <- rbind(mx[(minage+1):nrow(mx),],mx.fill) # or is it
    mx <- rbind(mx.fill,mx[(minage+1):nrow(mx),]) # I think this is right!
  }

  source("D:/Github/ICM/Scripts/functions/backwards_sim.R")

  tst <- icm.sim(years,
                 mat.age = age.mat,
                 nm = prop.nat.mort,
                 w.age = weight.age,
                 ages = minage:maxage,
                 rems = rem$rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 500,
                 sd.mat = 0.5,
                 sd.nm = 0.5,
                 sd.wt = 0.5,
                 sd.fecund = 0.5)

@@

@@

  nm = prop.nat.mort

  source("D:/Github/ICM/Scripts/functions/tuning_sim.R")
  N.end <- vpa.abund[length(vpa.abund)]
  tunes <- tune.sim(years,
                    n.steps = 100,
                    abund.ts = vpa.abund,
                    mat.age = age.mat,
                    nm = prop.nat.mort,
                    w.age = weight.age,
                    ages = minage:maxage,
                    rems = rem$rem,
                    fecund = mx,
                    N.end = N.end,
                    pop.model = 'exponential')

  tuned.res[[i]] <- tunes

@@

     # Difference in m's
   p5 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=diff.m),color='red') +
          xlab("") +
          ylab("Difference in natural mortality (Proportional)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p5, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/diff_m_", i, ".png"), base_width = 6,base_height = 4)
     # % Difference in m's
   p6 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=per.diff.m),color='red') +
          xlab("") +
          ylab("Difference in natural mortality (%)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p6, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/per_diff_m_", i, ".png"), base_width = 6,base_height = 4)

   p7 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=mn.m),color='blue') +
          geom_line(data=tunes$res[-1,], aes(x=year,y=vpa.m),color='black') +
          xlab("") +
          ylab("Natural mortality (Proportional)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          ggtitle(i) #+
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p7, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/Nat_M_", i, ".png"), base_width = 6,base_height = 4)

@@

}
saveRDS(object = tuned.res,file = "D:/Github/ICM/Results/tuned_fits.Rds")
saveRDS(object = back.res,file = "D:/Github/ICM/Results/backwards_sim_results.Rds")
saveRDS(object = back.ts,file = "D:/Github/ICM/Results/backwards_sim_ts_summaries.Rds")
saveRDS(object = back.ts.quants,file = "D:/Github/ICM/Results/backwards_sim_ts_quantiles_summaries.Rds")
```

>>>>

```{r, load-stuff,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# OK, so using the ICES assessments here's what we get for North Sea cod.
library(readxl)
library(tidyverse)
library(rio)
library(ggthemes)
library(cowplot)

source("D:/Github/ICM/Scripts/functions/backwards_sim.R")
source("D:/Github/ICM/Scripts/functions/tuning_sim.R")

# Choose 5 ICES stocks that we have the necessary data for
#ASR <- read_xlsx("../Data/ASR_2018.xlsx", sheet = "ICES")
ASR1 <- read.csv("D:/Github/ICM/Data/ICM_data_NE_atlantic_stocks.csv")
ASR <- read.csv("D:/Github/ICM/Data/ICM_data_NW_atlantic_stocks.csv")

ASR <- rbind(ASR1,ASR)
datatypes <- unique(gsub(x = names(ASR), pattern = "[^a-zA-Z]", replacement=""))

# Replace 0's in Num.tot with NA so the rest of this works as Freya designed it to...
ASR$Num.tot[ASR$Num.tot == 0] <- NA
# we want:
# Year, Num, WA, Catch, AM, NM, StockID, Management, Area, Order, Family, Genus, Species
ASRdat <- ASR[,c(grep(x=names(ASR), "Num"),
                 grep(x=names(ASR), "WA"),
                 grep(x=names(ASR), "Catch"),
                 grep(x=names(ASR), "AM"),
                 grep(x=names(ASR), "NM"))]
# Making all the data numeric that should be numeric
ASRdat <- apply(X = ASRdat, 2, as.numeric)
# Getting the species info back
ASRsp <- ASR[, which(!1:length(names(ASR)) %in% grep(x=names(ASR), ".", fixed=T))]
# And binding it all back together
ASR_trim <- cbind(ASRsp, ASRdat)
# need a unique ID for stock
#table(ASR_trim$Management, ASR_trim$Species)
ASR_trim$Stock <- paste0(ASR_trim$Management, "_", ASR_trim$Area, "_", ASR_trim$Genus, "_", ASR_trim$Species)

@@

```

```{r, prep-data,echo=F, message=F, warning=F, fig.width=6, fig.height=4}

@@

# Stocks <- ASR_stocks %>%
#                       dplyr::group_by(Stock, Species, type) %>%
#                       dplyr::summarize(ages=length(unique(age)),
#                                 years=length(unique(Year))) %>%
#                       dplyr::arrange(-years, -ages) %>%
#                       dplyr::filter(!Species=="morhua") %>%
#                       dplyr::distinct(Stock) %>%
#                       dplyr::pull(Stock)
# 
# print(Stocks)
Stocks <- Stocks$Stock

# Going to remove Sebastes norvegicus because we can't get fecundity as the only data we have is total numbers.
#Stocks <- Stocks[Stocks != "ICES-AFWG_DEEP1-2_Sebastes_norvegicus"]
# Esmarki doesn't work great as it is assessed using a quarterly modle.
#Stocks <- Stocks[Stocks != "ICES-WGNSSK_NS 4-3aN_Trisopterus_esmarkii"]

years.tmp <- NULL
pnm.tmp <- NULL
waa.tmp <- NULL
ages.tmp <- NULL
rem.tmp <- NULL
mx.tmp <- NULL
NE.tmp <- NULL
vpa.tmp <- NULL
am.tmp <- NULL
mr.tmp <- NULL
for(i in Stocks)
{
  print(i)
  ASR_sub <- ASR_long %>%
    dplyr::select(Year, Stock, type, age, value) %>%
    dplyr::filter(Stock==i)

@@

  # For the years of mx without SSB info to inform them
  if(minage >0 ) 
  {
    mx.fill <- as.data.frame(matrix(rep(colMeans(mx,na.rm=T),minage),nrow=minage,byrow=T),colnames = names(mx))
    names(mx) <- names(mx.fill)
   # mx <- rbind(mx[(minage+1):nrow(mx),],mx.fill) # or is it
    mx <- rbind(mx.fill,mx[(minage+1):nrow(mx),]) # I think this is right!
  }
  years.tmp[[i]] <- years
  pnm.tmp[[i]] <- prop.nat.mort
  waa.tmp[[i]] <- weight.age
  ages.tmp[[i]] <- minage:maxage
  rem.tmp[[i]] <- rem
  mx.tmp[[i]] <- mx
  NE.tmp[[i]] <- N.end
  vpa.tmp[[i]] <- vpa.abund
  am.tmp[[i]] <- age.mat
  mr.tmp[[i]] <- missing_rem
} #end input data loop

save(years.tmp,pnm.tmp,waa.tmp,ages.tmp,rem.tmp,mx.tmp,NE.tmp,vpa.tmp,mr.tmp,am.tmp,Stocks,ASR_stocks,ASR_long,file = "D:/Github/ICM/Results/model_inputs.Rdata")

```

```{r, run-mods,echo=F, message=F, warning=F}

load(file = "D:/Github/ICM/Results/model_inputs.Rdata")

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]

  tst <- icm.sim(years,
                 mat.age = age.mat,
                 nm = prop.nat.mort,
                 w.age = weight.age,
                 ages = ages,
                 rems = rem$rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 500,
                 sd.mat = 0.5,
                 sd.nm = 0.5,
                 sd.wt = 0.5,
                 sd.fecund = 0.5)

@@

  N.end <- vpa.abund[length(vpa.abund)]
  tunes <- tune.sim(years,
                    n.steps = 100,
                    abund.ts = vpa.abund,
                    mat.age = age.mat,
                    nm = prop.nat.mort,
                    w.age = weight.age,
                    ages = ages,
                    rems = rem$rem,
                    fecund = mx,
                    N.end = N.end,
                    pop.model = 'exponential')

  tuned.res[[i]] <- tunes
} # end i in Stocks look
# Save all these
saveRDS(object = tuned.res,file = "D:/Github/ICM/Results/tuned_fits.Rds")
saveRDS(object = back.res,file = "D:/Github/ICM/Results/backwards_sim_results.Rds")
saveRDS(object = back.ts,file = "D:/Github/ICM/Results/backwards_sim_ts_summaries.Rds")
saveRDS(object = back.ts.quants,file = "D:/Github/ICM/Results/backwards_sim_ts_quantiles_summaries.Rds")
```

```{r, make-figs,echo=F, message=F, warning=F, fig.width=6, fig.height=4}

# Now we can run the figures separately.
tuned.res <- readRDS(file = "D:/Github/ICM/Results/tuned_fits.Rds")
back.ts <- readRDS(file = "D:/Github/ICM/Results/backwards_sim_ts_summaries.Rds")
back.ts.quants <- readRDS(file = "D:/Github/ICM/Results/backwards_sim_ts_quantiles_summaries.Rds")
back.res <- readRDS(file = "D:/Github/ICM/Results/backwards_sim_results.Rds")
load(file = "D:/Github/ICM/Results/model_inputs.Rdata")

for(i in Stocks)
{
  # Unpack the lists
  tunes <- tuned.res[[i]] 
  did.it.work <- back.ts[[i]] 
  quants <- back.ts.quants[[i]] 
  missing_rem <- mr.tmp[[i]] 
  tst <- back.res[[i]]

@@

@@

     # Difference in m's
   p5 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=diff.m),color='red') +
          xlab("") +
          ylab("Difference in natural mortality (Instantaneous)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p5, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/diff_m_", i, ".png"), base_width = 6,base_height = 4)
     # % Difference in m's
   p6 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=per.diff.m),color='red') +
          xlab("") +
          ylab("Difference in natural mortality (%)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p6, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/per_diff_m_", i, ".png"), base_width = 6,base_height = 4)

   p7 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=mn.m),color='blue') +
          geom_line(data=tunes$res[-1,], aes(x=year,y=vpa.m),color='black') +
          xlab("") +
          ylab("Natural mortality (Instantaneous)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          ggtitle(i) #+
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p7, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/Nat_M_", i, ".png"), base_width = 6,base_height = 4)

@@

}

```
